---
title: "Project_Work3"
author: "Noah McIntire (nem2pq)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import Packages
```{r}
library(dplyr) # for data wrangling
library(lubridate) # date manipulation
library(TSstudio) # time series interactive viz
library(forecast) # time series library
library(tseries) # for adf.test
```


# Load in data and drop unneccesary columns and rows
```{r}
cville_weather <- read.csv("3291089 (1).csv")

cville_weather <- subset(cville_weather, select = -c(DAPR, MDPR, WESD))
cville_weather <- cville_weather[-c(1,2),]
head(cville_weather)
```
Split into training and test data, make sure date ranges are correct
```{r}
climate_train <- cville_weather[1:4353,]
head(climate_train)
climate_test <- cville_weather[4354:5448,]
head(climate_test)
range(climate_train$DATE)
range(climate_test$DATE)
```
# Check for NAs in the dataset, and drop them if present
```{r}
anyNA(climate_train)
anyNA(climate_test)
climate_train<-climate_train[complete.cases(climate_train), ]
climate_test<-climate_test[complete.cases(climate_test), ]
anyNA(climate_train)
anyNA(climate_test)
```

# Preprocess data for High temperature training
```{r}
train_high <- climate_train %>% select(DATE, TMAX)
test_high <- climate_test %>% select(DATE, TMAX)
```

```{r}
ts_httrain <- ts(data = train_high$TMAX,
                start = c(2008,05),
                frequency = 365)
autoplot(ts_httrain)
```

```{r}
ts_httest <- ts(data = test_high$TMAX,
                start = c(2020,04),
                frequency = 365)
autoplot(ts_httest)
```

#Decompse time series into a new set
```{r}
climate_cvht <- decompose(ts_httrain)
plot(climate_cvht)
```


# Creating a holt winters model
```{r}
climate_hwht <- HoltWinters(ts_httrain,seasonal = "additive")
climate_hwht
```


```{r}
climate_forecastht <- forecast(climate_hwht, h=1090)
climate_forecastht
```


```{r}
forecast::accuracy(climate_forecastht, ts_httest)
```

# Repeat with low temp, obs temp?
# Preprocess data for High temperature training
```{r}
train_low <- climate_train %>% select(DATE, TMIN)
test_low <- climate_test %>% select(DATE, TMIN)
```

```{r}
ts_lttrain <- ts(data = train_low$TMIN,
                start = c(2008,05),
                frequency = 365)
autoplot(ts_lttrain)
```

```{r}
ts_lttest <- ts(data = test_low$TMIN,
                start = c(2020,04),
                frequency = 365)
autoplot(ts_lttest)
```

#Decompse time series into a new set
```{r}
climate_cvlt <- decompose(ts_lttrain)
plot(climate_cvlt)
```


# Creating a holt winters model
```{r}
climate_hwlt <- HoltWinters(ts_lttrain,seasonal = "additive")
climate_hwlt
```


```{r}
climate_forecastlt <- forecast(climate_hwlt, h=1090)
climate_forecastlt
```


```{r}
forecast::accuracy(climate_forecastlt, ts_lttest)
```


# Now the Observed Temperature


# Preprocess data for High temperature training
```{r}
train_ob <- climate_train %>% select(DATE, TOBS)
test_ob <- climate_test %>% select(DATE, TOBS)
```

```{r}
ts_obtrain <- ts(data = train_ob$TOBS,
                start = c(2008,05),
                frequency = 365)
autoplot(ts_obtrain)
```

```{r}
ts_obtest <- ts(data = test_ob$TOBS,
                start = c(2020,04),
                frequency = 365)
autoplot(ts_obtest)
```

#Decompse time series into a new set
```{r}
climate_cvob <- decompose(ts_obtrain)
plot(climate_cvob)
```


# Creating a holt winters model
```{r}
climate_hwob <- HoltWinters(ts_obtrain,seasonal = "additive")
climate_hwob
```


```{r}
climate_forecastob <- forecast(climate_hwob, h=1090)
climate_forecastob
```


```{r}
forecast::accuracy(climate_forecastob, ts_obtest)
```

# Now for precipitation
# Preprocess data for High temperature training
```{r}
train_high <- climate_train %>% select(DATE, PRCP)
test_high <- climate_test %>% select(DATE, PRCP)
```

```{r}
ts_pctrain <- ts(data = train_high$PRCP,
                start = c(2008,05),
                frequency = 365)
autoplot(ts_pctrain)
```

```{r}
ts_pctest <- ts(data = test_high$PRCP,
                start = c(2020,04),
                frequency = 365)
autoplot(ts_pctest)
```

#Decompse time series into a new set
```{r}
climate_cvpc <- decompose(ts_pctrain)
plot(climate_cvpc)
```


# Creating a holt winters model
```{r}
climate_hwpc <- HoltWinters(ts_pctrain,seasonal = "additive")
climate_hwpc
```


```{r}
climate_forecastpc <- forecast(climate_hwpc, h=1090)
climate_forecastpc
```


```{r}
forecast::accuracy(climate_forecastpc, ts_pctest)
```



